---
title: "STAT506PS4"
format:
  pdf:
    documentclass: scrreprt
    toc: true
    toc-depth: 3
    listings: true
editor: visual
execute:
  error: TRUE
---

```{r}
library("tidyverse")
library("nycflights13")
library("dplyr")
library("survey")
library("haven")
```

## Question1

\(1\)

```{r}
data("flights")
```

```{r}
colnames(flights)
```

```{r}
#The first table Dep_delay
Dep_delay <- flights%>%
  group_by(origin)%>%
  dplyr::filter(n() >= 10) %>%   
  summarize(Dep_delay_mean = mean(dep_delay,na.rm = TRUE), Dep_delay_median = median(dep_delay,na.rm = TRUE))%>%
  arrange(desc(Dep_delay_mean))%>%
  left_join(airports, by = c("origin" = "faa")) %>%
  select(origin = name, Dep_delay_mean, Dep_delay_median)
print(Dep_delay,n = Inf)
```

```{r}
#The second table Arr_delay
Arr_delay <- flights %>%
  group_by(dest)%>%
    dplyr::filter(n()>=10)%>%
  summarize(Arr_delay_mean = mean(arr_delay,na.rm = TRUE),Arr_delay_median = median(arr_delay,na.rm = TRUE))%>%
  arrange(desc(Arr_delay_mean))%>%
  left_join(airports,by = c("dest" = "faa"))%>%
  select(dest = name,Arr_delay_mean, Arr_delay_median)
print(Arr_delay,n = Inf)
```

\(b\)

```{r}

```

## Question2

```{r}
get_temp <- function(month,year,data,celsius = FALSE,average_fn = mean){
  #Check valid year
  if (!is.numeric(year)|| year < min(data$year) || year > max(data$year)){
    stop("Invalid year input")
  }
  #Check valid year and transfer year to abbreviated month format
  month_names_abbrev <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec") 
  month_names_full <- c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")
  if (month %in% month_names_full){
  month <- month_names_abbrev[which(month_names_full == month)]
  } else if (is.numeric(month)){
  if(month > 12 || month < 1){
    stop("Invalid month input.")
  }
  month <- month_names_abbrev[month]
} else if(!month %in% month_names_abbrev) {
  stop("Invalid month")
}
  #get the result
  avg_temp <- data %>%
    filter(year == !!year & month == !!month) %>%
    summarise(avg_temperature = average_fn(temp)) %>%
    pull(avg_temperature)
  #check the format for temperature
  if(celsius) {
    avg_temp <- (avg_temp - 32) * 5/9
  }
  return(avg_temp)
}
```

```{r}
nnmaps <- read.csv("/Users/meredithwang/Downloads/chicago-nmmaps.csv")
get_temp("Apr", 1999, data = nnmaps)
```

```{r}
get_temp("Apr", 1999, data = nnmaps, celsius = TRUE)
```

```{r}
get_temp(10, 1998, data = nnmaps, average_fn = median)
```

```{r}
get_temp(13, 1998, data = nnmaps)
```

```{r}
get_temp(2, 2005, data = nnmaps)
```

```{r}
get_temp("November", 1999, data =nnmaps, celsius = TRUE,
         average_fn = function(x) {
           x %>% sort -> x
           x[2:(length(x) - 1)] %>% mean %>% return
         })
```

## Question3

\(a\)

``` sas
/* import data */
proc import datafile="/home/u63646271/sasuser.v94/recs2020_public_v5.csv" out=recs2020 replace;

/* Calculate the weights for all states */
proc sql noprint;
   /* weight of each state */
   create table state_weights as
   select state_name, sum(NWEIGHT) as total_weight
   from recs2020
   group by state_name;
   
   select sum(NWEIGHT) into :totalWeight from recs2020;
   
   /* percentage of each state and sort */
   create table state_percent as
   select state_name, 
          total_weight / &totalWeight as percentage
   from state_weights
   order by percentage desc;
run;

/* Calculate the highest percentage */
proc print data=state_percent(obs=1); 
   var state_name percentage;
run;

/* Calculate the Percentage of All Records for Michigan */
proc sql;
   select state_name, percentage 
   from state_percent 
   where state_name = "Michigan";
run;
```

\(b\)

``` sas
data positive_elec_cost;
    set recs2020;
    if DOLLAREL > 0;
run;

proc sgplot data=positive_elec_cost;
    histogram DOLLAREL;
    title "Histogram of Total Electricity Cost (Strictly Positive Values)";
    xaxis label="Total Electricity Cost";
    yaxis label="Number of Houses";
run;
```

\(c\)

``` sas
data log_elec_cost;
    set recs2020;
    if DOLLAREL > 0 then log_elec = log(DOLLAREL); 
    else delete; 
run;

/* Plot the histogram */
proc sgplot data=log_elec_cost;
    histogram log_elec;
    title "Histogram of Log-transformed Total Electricity Cost";
    xaxis label="Log of Total Electricity Cost";
    yaxis label="Number of Houses";
run;
```

\(d\)

``` sas
data prep_data;
    set recs2020;
    if DOLLAREL > 0 then log_elec = log(DOLLAREL);
    else delete; 

    if PRKGPLC1 = 1 then has_garage = 1;
    else has_garage = 0;
    
run;

proc reg data=prep_data;
    model log_elec = TOTROOMS has_garage;
    weight NWEIGHT; 
    title "Linear Regression Model Predicting Log of Electricity Cost";
run;
```

（e)

``` sas
/* Fit the model and save the predicted values */
proc reg data=prep_data outest=estimates noprint;
    model log_elec = TOTROOMS has_garage;
    output out=predicted_data p=pred_log_elec;
run;

data predicted_data;
    set predicted_data;
    pred_elec_cost = exp(pred_log_elec); 
run;

proc sgplot data=predicted_data;
    scatter x=DOLLAREL y=pred_elec_cost;
    lineparm x=0 y=0 slope=1 / clip; 
    title "Scatterplot of Predicted vs. Actual Total Electricity Cost";
run;
```

## Question4

\(b\)

``` sas
/* import data to SAS */
proc import datafile="/home/u63646271/sasuser.v94/public2022.csv" out=public2022 replace;

/* Question4(b) Choose variables of interest */
proc sql;
    create table select_data as
    select CaseID, weight_pop,
        B3, 
        ND2, 
        B7_b, 
        GH1, 
        educ_4cat, 
        ppethm
    from public2022;
run;
```

\(c\)

``` sas
proc export
   data = select_data
   outfile = "/home/u63646271/sasuser.v94/analysis_data.csv"
   dbms = CSV;
run;
```

\(d\)

``` stata
 import delimited "/Users/meredithwang/Downloads/analysis_data.csv"
(encoding automatically selected: ISO-8859-1)
(8 vars, 11,667 obs)
```

After I import the data into STATA, the data has 6 variables as we selected before and also 11668 observations which is also the same as the Cook book.

\(e\)

``` stata
gen wealth_binary = (b3 == "Somewhat worse off" | b3 == "Much worse off")
encode nd2, generate(nd2_num)
label define nd2_label 1 "Much higher" 2 "Somewhat higher" 3 "About the same" 4 "Somewhat lower" 5 "Much lower"
label values nd2_num nd2_label
encode b7_b, generate(b7_b_num)
label define b7_b_label 1 "Poor" 2 "Only fair" 3 "Good" 4 "Excellent"
label values b7_b_num b7_b_label
encode gh1, generate(gh1_num)
label define gh1_label 1 "Own your home with a mortgage or loan" 2 "Own your home free and clear (without a mortgage or loan)" 3 "Pay rent" 4 "Neither own nor pay rent"
label values gh1_num gh1_label
encode educ_4cat, generate(educ_4cat_num)
label define educ_4cat_label 1 "Less than a high school degree" 2 "High school degree or GED" 3 "Some college/technical or associates degree" 4 "Bachelor's degree or more"
label values educ_4cat_num educ_4cat_label
encode ppethm, generate(ppethm_num)
label define ppethm_label 1 "White, Non-Hispanic" 2 "Black, Non-Hispanic" 3 "Other, Non-Hispanic" 4 "Hispanic" 5 "2+ Races, Non-Hispanic"
label values ppethm_num ppethm_label
```

``` stata
. svyset caseid [pw=weight_pop]
. svy: logistic wealth_binary i.nd2_num i.b7_b_num i.gh1_num i.educ_4cat_num i.ppethm_num
(running logistic on estimation sample)

Survey: Logistic regression

Number of strata =      1                        Number of obs   =      11,667
Number of PSUs   = 11,667                        Population size = 255,114,223
                                                 Design df       =      11,666
                                                 F(17, 11650)    =       57.42
                                                 Prob > F        =      0.0000

----------------------------------------------------------------------------------------------------------------------------
                                                           |             Linearized
                                             wealth_binary | Odds ratio   std. err.      t    P>|t|     [95% conf. interval]
-----------------------------------------------------------+----------------------------------------------------------------
                                                   nd2_num |
                                          Somewhat higher  |   1.074973   .0909575     0.85   0.393     .9106824    1.268903
                                           About the same  |   .8474572   .1282401    -1.09   0.274     .6299379    1.140087
                                           Somewhat lower  |   .9813484   .0540987    -0.34   0.733     .8808343    1.093332
                                               Much lower  |   .8243532   .1579435    -1.01   0.313     .5662509    1.200101
                                                           |
                                                  b7_b_num |
                                                Only fair  |   1.974776   .6917774     1.94   0.052      .993814    3.924013
                                                     Good  |   3.953395   1.362284     3.99   0.000     2.011999    7.768062
                                                Excellent  |   11.98335   4.130326     7.21   0.000     6.097651    23.55017
                                                           |
                                                   gh1_num |
Own your home free and clear (without a mortgage or loan)  |   1.517157   .1572363     4.02   0.000     1.238237    1.858905
                                                 Pay rent  |   1.422295   .1417282     3.54   0.000     1.169932    1.729094
                                 Neither own nor pay rent  |   1.385568   .1436771     3.14   0.002     1.130715    1.697864
                                                           |
                                             educ_4cat_num |
                                High school degree or GED  |   1.226953   .0770309     3.26   0.001     1.084881    1.387631
              Some college/technical or associates degree  |   1.396686    .155816     2.99   0.003     1.122349     1.73808
                                Bachelor's degree or more  |   1.162817    .061381     2.86   0.004     1.048516    1.289579
                                                           |
                                                ppethm_num |
                                      Black, Non-Hispanic  |   .4458974   .0687591    -5.24   0.000     .3295815    .6032635
                                      Other, Non-Hispanic  |   .7637862   .1143198    -1.80   0.072     .5695802    1.024209
                                                 Hispanic  |    .635055   .1135432    -2.54   0.011     .4473082    .9016039
                                   2+ Races, Non-Hispanic  |   .9073112   .1243138    -0.71   0.478     .6936145    1.186846
                                                           |
                                                     _cons |    .076848   .0289352    -6.81   0.000     .0367368    .1607549
----------------------------------------------------------------------------------------------------------------------------
Note: _cons estimates baseline odds.
```

Based on the result above, we can find that none of the categorical of predictor 'nd2_num' has p-value less than 0.05, which gives us that none of the categories of the predictorare significantly different from the reference category in terms of predicting financial situation relative to the previous year.

\(f\)

``` stata
.save "/Users/meredithwang/Desktop/Course/STAT506/Umich_stats506/analysis_data2.dta"
```

\(g\)

```{r}
public_2022 <- read_dta("/Users/meredithwang/Desktop/Course/STAT506/Umich_stats506/analysis_data2.dta")
names(public_2022)
```

```{r}
des <- svydesign(id = ~caseid, weights = ~weight_pop, data = public_2022)
```

```{r}
model <- svyglm(wealth_binary ~ factor(nd2_num) + factor(b7_b_num) + factor(gh1_num) + 
                factor(educ_4cat_num) + factor(ppethm_num), 
                design = des, family = quasibinomial())
```

```{r}
pseudo_R2 <- 1 - (model$deviance / model$null.deviance)
print(pseudo_R2)
```

The pseudo R\^2 provides a measure of how well the predictors explain the variability in the response variable. Since this R\^2 is low in this case, we don't consider it as a good fit for this model.
